Entropy-SGD: Biasing Gradient Descent Into Wide Valleys

### 2016-12-20 doc/rebuttal.txt

We are happy to provide clarifications to the questions posed. We enclose them below.

"mostly reinventing SVI"
SVI computes the posterior distribution of hidden variables given data by maximizing the ELBO. Please note that we do not have hidden variables in our formulation; "x" are the variables of our objective f(x), not data. We are interested in maximizing the free energy F(x,\gamma) as defined in Eqn. 4. Simply speaking, this is the *free energy of the energy landscape* (defined via a Gibbs distribution). This is thus unrelated to marginal likelihood or variational ELBO where one can impose entropic constraints on the posterior distribution of the *hidden variables* via a prior.

"I suspect it'd become clear that you are maximizing the marginal likelihood"
Please note that we are not maximizing the marginal likelihood.

"how is this better than SGLD", "bizarre that you compare against Adam instead of SGLD"
SGLD is an MCMC algorithm that draws samples from a given distribution. If the step-size goes to zero slowly enough, akin to all MCMC algorithms, it converges to the maximum of the likelihood in exponentially long time-scales (see algorithms like SGHMC [CFG14], Santa [CCG15] etc. that train Bayesian neural networks using MCMC algorithms). Note that SGLD does not optimize the marginal likelihood because there is no notion of hidden variables in vanilla SGLD. We provide a brief review of SGLD in Appendix A of our paper.

We do not know of any results in literature that train large deep networks such as the one used for CIFAR to get competitive error rates using SGLD. In our experience, vanilla SGLD performs poorly due to non-energy preserving dynamics while more complex algorithms like [CFG14 or CCF15] have too many sensitive hyper-parameters. We would like to emphasize that Entropy-SGD simply uses MCMC sampling to estimate Eqn. 7, it is unrelated to SGLD otherwise.

On the other hand, Adam is an algorithm for computing the maximum of a the likelihood of data given parameters or (equivalently, minimize the loss function). Entropy-SGD is an algorithm designed for minimizing the loss function f(x), in particular, it is not an MCMC algorithm that draws samples from the likelihood. It however does not explicitly do so and instead maximizes the local entropy. We therefore compare Entropy-SGD with state-of-the-art algorithms for training deep networks like Adam and SGD.

"frustrating that you discuss free energy and entropy without precise definitions"
Local entropy (local free energy) is formally defined in Def. 1 (Eqn. 4) but it is introduced as early as pg. 2 in the Introduction. The discussion towards the end of Sec. 3 (pg. 5 first line) defines the classical entropy of a Gibbs distribution. The beginning of Sec. 3 based on Fig. 2 is intended to explain things to the reader at an intuitive level before proceeding to the formal definitions which involve a little bit of jargon.

Free energy and classical entropy are related by the informal description "free energy = internal energy - temperature x entropy". Formally, the relation is
    F(\beta) = U(\beta) - \frac{1}{\beta} S(\beta)
where F(\beta) is the log-partition function (free energy) defined as -\beta^{-1} \log Z(\beta), U(\beta) = \partial_\beta (\beta F(\beta)) is the internal energy and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.

Note that the above equation is about the entirety of the Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn. 3. We can add this discussion to the paper.

### 2016-12-22 doc/rebuttal.txt

We are happy to provide clarifications to the questions posed. We enclose them below.

"mostly reinventing SVI"
SVI computes the posterior distribution of hidden variables given data by maximizing the ELBO. Please note that we do not have hidden variables in our formulation; "x" are the variables of our objective f(x), not data. We are interested in maximizing the free energy F(x,\gamma) as defined in Eqn. 4. This is the *free energy of the energy landscape* (defined via a Gibbs distribution). This is thus unrelated to marginal likelihood or variational ELBO where one can impose entropic constraints on the posterior distribution of the *hidden variables* via a prior.

"I suspect it'd become clear that you are maximizing the marginal likelihood"
Please note that we are not maximizing the marginal likelihood, as discussed above.

@@

"how is this better than SGLD", "bizarre that you compare against Adam instead of SGLD"
SGLD is an MCMC algorithm that draws samples from a given distribution. If the step-size goes to zero slowly enough, akin to all MCMC algorithms, it converges to the maximum of the likelihood in exponentially long time-scales (see algorithms like SGHMC [CFG14], Santa [CCG15] etc. that train Bayesian neural networks using MCMC algorithms). Note that SGLD does not optimize the marginal likelihood because there is no notion of hidden variables in vanilla SGLD. We provide a brief review of SGLD in Appendix A of our paper.

We do not know of any results in the literature that train large deep networks such as the one used for CIFAR to get competitive error rates using SGLD. We would like to emphasize that Entropy-SGD simply uses MCMC sampling to estimate Eqn. 7 but it is unrelated to SGLD otherwise.

@@

On the other hand, Adam is an algorithm for computing the maximum of a the likelihood of data given parameters or (equivalently, minimize the loss function). Entropy-SGD is an algorithm designed for minimizing the loss function f(x), in particular, it is not an MCMC algorithm that draws samples from the likelihood. It however does not explicitly do so and instead maximizes the local entropy. We therefore compare Entropy-SGD with state-of-the-art algorithms for training deep networks like Adam and SGD.

"frustrating that you discuss free energy and entropy without precise definitions"
Local entropy (local free energy) is formally defined in Def. 1 (Eqn. 4) but it is already introduced on pg. 2 in the Introduction. The discussion towards the end of Sec. 3 (pg. 5, first line) defines the classical entropy of a Gibbs distribution. The beginning of Sec. 3 based on Fig. 2 is intended to explain things to the reader at an intuitive level before proceeding to the formal definitions which are nevertheless present.

Free energy and classical entropy are related by the informal description "free energy = internal energy - temperature x entropy". Formally, the relation is
    F(\beta) = U(\beta) - \frac{1}{\beta} S(\beta)
where the log-partition function (free energy) is defined as F(\beta) = -\beta^{-1} \log Z(\beta), the internal energy is U(\beta) = \partial_\beta (\beta F(\beta)) and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.

Note that the above equation is about the entire Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn. 3. We can add this discussion to the paper.

### 2016-12-22 doc/rebuttal.txt

Free energy and classical entropy are related by the informal description "free energy = internal energy - temperature x entropy". Formally, the relation is
    F(\beta) = U(\beta) - \frac{1}{\beta} S(\beta)
where the log-partition function (free energy) is defined as F(\beta) = -\beta^{-1} \log Z(\beta), the internal energy is U(\beta) = \partial_\beta (\beta F(\beta)) and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.

Note that the above equation is about the entire Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn. 3. We can add this discussion to the paper.

[CFG14] Chen, Tianqi, Emily B. Fox, and Carlos Guestrin. "Stochastic Gradient Hamiltonian Monte Carlo." ICML. 2014.
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).

----------------------------------------------------------------------

Thanks for your comments, allow us to elaborate further.

In the notation of the our paper let x denote the parameters and \xi_k denote individual samples. The ELBO is then given by
    L = E_{x \sim q(x)} [\log p(\Xi | x)] - KL(q(x) || p(x))
where p(x) is a fixed prior on the parameters and q(x) is the posterior distribution q(x | \Xi). In mean-field variational inference, we make a mean-field approximation of the distribution q(x) and have to come up with a prior p(x) (see [GG15,BKM16]). Using a Monte-Carlo approximation, the gradient of ELBO can be given by
    dL = \sum_{k=1}^N \log p(\xi | x_k) - KL(q(x) || p(x))
for samples x_k \sim q(x). The first term is simply how well x fits the data while the second term is the distance of the posterior from the prior. If p(x) = N(0, I), one can evaluate the KL term in closed form if q(x) is also a product of Gaussians etc. and compute dL. Note that this approach requires a number of critical assumptions such as a factorizing of q(x) for ease of sampling in the Monte-Carlo term and evaluating the KL term, Gaussian / some other easy to compute prior on weights. These assumptions are not typically valid in CNNs.

The gradient in Eqn. 7 in our paper can be written as:
    dF = -\gamma (x - E_{x' \sim r(x'; x, \xi_k)} [x'])
with the corresponding Monte-Carlo estimate of the expectation. The crucial difference is that the distribution r(x'; x) in our case is the one described in Eqn. 3, i.e.,
    \log r(x'; x) = \log q(x')  + \gamma |x-x'|^2/2 - \log Z_{x,\gamma}.
q(x') can be the same distribution as variational inference above. However, note that r(x'; x) in our work is a function of the current iterate x itself. This is akin to a "prior" that moves with the iterates x which is very different from a Bayesian neural network with a fixed prior.

It is thus easy to see the differences between a variational approach on the weights and our method. We agree that the motivation for our work and approximate inference methods is similar but our method is much different from variational approximation. We'd be very interested if the reviewer has other ways of connecting our paper with it.

"SGLD is a simpler method"
We respectfully disagree. In practice, it is very hard to get SGLD to work as well as SGD or Adam. This has motivated numerous other MCMC algorithms such as SGHMC, SGNHT [DFB14], Santa [CCF15] etc. None of these have yet been shown to beat SGD's error rates for large deep networks.

While it is not hard to add experiments comparing Entropy-SGD to SGLD, this would be a digression. The error rates for baselines in our paper are close to the best error rates reported in the literature for these networks, e.g., on CIFAR, we get 8.3% with SGD and 8.65% with Entropy-SGD both of which are better than those of the original authors (9.08%) [SDB15]. The authors in [CCF15] get a comparable error rate on MNIST (0.47% vs 0.48% for us) with a larger network (Table 1) and a much more complex algorithm. [CCF15] also provides the results for SGLD on LeNet which are much worse than ours. Therefore, we feel our comparisons against baselines are entirely fair.

[GG15] Gal, Yarin, and Zoubin Ghahramani. "Dropout as a Bayesian approximation: Representing model uncertainty in deep learning." arXiv:1506.02142 (2015).
[BKM16] Blei, David M., Alp Kucukelbir, and Jon D. McAuliffe. "Variational inference: A review for statisticians." arXiv:1601.00670 (2016).
[SDB15] Springenberg, Jost Tobias, et al. "Striving for simplicity: The all convolutional net." arXiv:1412.6806 (2014).
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).
[DFB14] Ding, Nan, et al. "Bayesian sampling using stochastic gradient thermostats." NIPS. 2014.

### 2016-12-29 doc/rebuttal.txt

Free energy and classical entropy are related by the informal description "free energy = internal energy - temperature x entropy". Formally, the relation is
    F(\beta) = U(\beta) - \frac{1}{\beta} S(\beta)
where the log-partition function (free energy) is defined as F(\beta) = -\beta^{-1} \log Z(\beta), the internal energy is U(\beta) = \partial_\beta (\beta F(\beta)) and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.

Note that the above equation is about the entire Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn. 3. We can add this discussion to the paper.

[CFG14] Chen, Tianqi, Emily B. Fox, and Carlos Guestrin. "Stochastic Gradient Hamiltonian Monte Carlo." ICML. 2014.
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).

----------------------------------------------------------------------

Thanks for your comments, allow us to elaborate.

Let \Xi denote the dataset, z denote the weights and x be the parameters of the variational distribution on weights, i.e., the weights are sampled from the distribution q_x(z) := p_x(z | \Xi). Define the loss as the log-likehood of data, viz. -f(z) := \log p(\Xi | z). The ELBO is then given by
(1) \log p(\Xi)
        \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z))
and we typically maximize the right hand side wrt x. Let us emphasize that in variational inference, x are the parameters and z are random variables. Also p(z) is a fixed (parameter-free) prior on the weights z.

As noted in our previous comment, while our approach is not variational in nature, we shall now aim to connect the ELBO with local entropy. Focus on the local entropy term defined in Eqn. 4 in the paper: we have
(2) \log F(x,\gamma)
        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz
        \geq \int_{z \in Z} [-f(z; \Xi) - \gamma/2 |x-z|^2] dz.
Eqns. (1) and (2) are thus very different objects. In general, there does not exist a direct mapping between local entropy and variational inference.

We can however provide the following connected for a special case. Consider the family of flat variational distributions q_x(z) such that the distribution is uniform on some compact set around x, e.g., q_x(z) \propto c if |x-z| < C and zero otherwise. We will pick a special non-fixed prior p(z) \propto \exp(-\gamma/2 |x-z|^2). The expectation in Eqn. (1) can now be simplified as:
(3) \log p(\Xi)
        \geq \int_{z \in Z} [-f(z)] - KL(q_x(z) || p(z))
        = -\int_{z \in Z} c*f(z) dz + H(q_x(z)) - \int_{z \in Z} c*\gamma/2 |x-z|^2 dz;
The differential entropy entropy H(q_x(z)) is a constant wrt x. The other two terms are the same as Eqn. (2) (multiplied by a constant c).

In other words "local entropy is similar to performing variational inference within a special class of flat distributions around the paramter x". Such an explanation also validates our claim to finding flat regions in the energy landscape. The above mapping is however only heuristic because the prior p(z) is not fixed and depends on x itself which is unusual. The gradient of local entropy clarifies this:
    dF  = -\gamma (x - E_{x' \sim r(z; x)} [z]);
where the distribution r(z; x) is
    \log r(z; x) \propto \log p(\Xi | z)  - \gamma/2 |x-z|^2
and is thus a likelihood and prior that "moves" along with current iteratre x itself.

It is thus easy to see the differences between a variational approach on the weights and our method. At present we do not see a more rigorous way to reduce our method to SVI, but we would welcome specific suggestions if that is indeed the case.

We have tested SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and can add the results to the paper. In summary, after a hyper-parameter search, the best results we obtained yielded a test error of LeNet of 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs. Disregarding the slow convergence of SGLD, its generalization error is worse than the results in our paper, viz. 0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD). For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (and get 0.47% error with Santa). To our knowledge, there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.

-------------------------------------------------------------------------

You claim that "For modern deep networks with a few million weights, computing the Hessian-vector product as done in HS97 is prohibitive".  But in general the HVP can be computed for only twice the time cost as a gradient evaluation.  Can you elaborate on why this is expensive in your case?

The above comment is for HVP of a general deep network, not our algorithm. And indeed, you are right that HVP can be computed using forward differencing with two back-props [P94, CSP93]. However, as [P94] mentions, it is a

[CSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS 1993.
[P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.

- Bregman vs. KL
- On the use of stochastic hessian information in optimization methods for machine learning
- https://goo.gl/qxWzgr

### 2016-12-29 doc/rebuttal.txt

Free energy and classical entropy are related by the informal description "free energy = internal energy - temperature x entropy". Formally, the relation is
    F(\beta) = U(\beta) - \frac{1}{\beta} S(\beta)
where the log-partition function (free energy) is defined as F(\beta) = -\beta^{-1} \log Z(\beta), the internal energy is U(\beta) = \partial_\beta (\beta F(\beta)) and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.

Note that the above equation is about the entire Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn. 3. We can add this discussion to the paper.

[CFG14] Chen, Tianqi, Emily B. Fox, and Carlos Guestrin. "Stochastic Gradient Hamiltonian Monte Carlo." ICML. 2014.
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).

----------------------------------------------------------------------

Allow us to elaborate further.

Let \Xi denote the dataset, z denote the weights and x be the parameters of a variational distribution on weights denoted as q_x(z). Define the loss to be the negative log-likelihood of data, viz. -f(z) := \log p(\Xi | z). The ELBO is then given by
(1) \log p(\Xi)
        \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z));
we maximize the right hand side with respect to x. The distribution p(z) is a fixed (parameter-free) prior on the weights.

We shall now try to connect the ELBO with local entropy. In our paper, we maximize the local entropy (Eqn. 4 in the paper) or equivalently, its logarithm:
(2) \log F(x,\gamma)
        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz
        \geq \int_{z \in Z} [-f(z; \Xi) - \gamma/2 |x-z|^2] dz.
The inequality follows by an application of Jensen's inequality. As noted in our previous comment, our approach is not variational in nature and it is easy to see that Eqns. (1) and (2) are quite different. In general, there does not seem to be direct reduction from local entropy to variational inference.

We can however provide the following connection in a special case. Consider a family of flat variational distributions q_x(z) such that the distribution has a mean x and it is uniform on a compact set, for instance, let q_x(z) \propto c if |x-z| < C and zero otherwise. We will also pick a special prior p(z) \propto \exp(-\gamma/2 |x-z|^2). The expectation in Eqn. (1) can now be simplified as:
(3) \log p(\Xi)
        \geq \int_{z \in Z} [-f(z)] - KL(q_x(z) || p(z))
        = -\int_{z \in Z} c*f(z) dz + H(q_x(z)) - \int_{z \in Z} c*\gamma/2 |x-z|^2 dz;
The differential entropy H(q_x(z)) is a constant with respect to x. For the purposes of maximization, ELBO as given in Eqn. (3) is thus equivalent to local entropy (upto an additive constant).

In other words, "local entropy is similar to performing variational inference within a class of uniform distributions with mean x". Such an explanation also validates our motivation to find flat regions in the energy landscape. The above discussion is however not entirely rigorous because the prior p(z) is not fixed and depends on the parameters x themselves. The gradient of local entropy (Eqn. 7 in the paper) clarifies this point:
    dF  = -\gamma (x - E_{z \sim r(z; x)} [z]);
where the distribution r(z; x) is
    \log r(z; x) \propto \log p(\Xi | z)  - \gamma/2 |x-z|^2
and is thus a data likelihood term with a prior that "moves" along with current iterate x.

At present we do not see a more rigorous way to reduce our method to SVI but we would welcome specific suggestions if that is indeed the case. We will update the paper with the above discussion.

In response to your previous comments, we have tested SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and can add the following results to the paper. After a hyper-parameter search, the best results we obtained yielded a test error of LeNet of 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs. Disregarding the slow convergence of SGLD, its generalization error is worse than the results in our paper, viz. 0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD). For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (and get 0.47% error with Santa). To our knowledge, there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.

[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).
-------------------------------------------------------------------------

You claim that "For modern deep networks with a few million weights, computing the Hessian-vector product as done in HS97 is prohibitive".  But in general the HVP can be computed for only twice the time cost as a gradient evaluation.  Can you elaborate on why this is expensive in your case?

The above comment is for HVP of the loss function of a general deep network, not local entropy. Indeed, HVPs can be computed using forward differencing with two back-props. However, as [P94] mentions, such an approximation is highly susceptible to numerical errors --- especially in high dimensions [P94, M10]. Moreover, estimating HVP on a single mini-batch is not enough, as [CSP93] mention one typically has to average the HVP over many mini-batches; this is expensive. For instance, the authors in [CSP93] average over 10-20 batches for estimating the largest eigenvalue of the Hessian in a network with ~60,000 weights. [P94] offers an accurate, although more complex, algorithm to compute HVPs using the R-operator.

[CSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS 1993.
[P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.
[M10] Martens, James. "Deep learning via Hessian-free optimization." ICML. 2010.

- On the use of stochastic hessian information in optimization methods for machine learning
- https://goo.gl/qxWzgr
- http://icml2010.haifa.il.ibm.com/papers/458.pdf

### 2017-01-02 doc/rebuttal.txt

[CFG14] Chen, Tianqi, Emily B. Fox, and Carlos Guestrin. "Stochastic Gradient Hamiltonian Monte Carlo." ICML. 2014.
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).

----------------------------------------------------------------------

Allow us to elaborate further.

Let \Xi denote the dataset, z denote the weights and x be the parameters of a variational distribution on weights denoted as q_x(z). Define the loss to be the negative log-likelihood of data, viz. -f(z) := \log p(\Xi | z). The ELBO is then given by
(1) \log p(\Xi)
        \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z));
where the right hand side would be maximized with respect to x. The distribution p(z) is a fixed (parameter-free) prior on the weights.

We shall now try to connect the ELBO with local entropy. In our paper, we maximize the local entropy (Eqn. 4 in the paper) or equivalently, its logarithm:
(2) \log F(x,\gamma)
        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz
        \geq \int_{z \in Z} [-f(z; \Xi) - \gamma/2 |x-z|^2] dz.
The inequality follows by an application of Jensen's inequality. As noted in our previous comment, our approach is not variational in nature and indeed, Eqns. (1) and (2) are quite different. In general, there does not seem to be direct reduction from local entropy to variational inference.

We can however provide the following connection in a special case. Consider a family of flat variational distributions q_x(z) such that the distribution has a mean x and it is uniform on a compact set, for instance, let q_x(z) \propto c if |x-z| < C and zero otherwise. We will also pick a special prior p(z) \propto \exp(-\gamma/2 |x-z|^2). The expectation in Eqn. (2) can now be simplified as:
(3) \log p(\Xi)
        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z))
        = -\int_{z \in Z} c*f(z) dz + H(q_x(z)) - \int_{z \in Z} c*\gamma/2 |x-z|^2 dz;
The differential entropy H(q_x(z)) is a constant with respect to x. For the purposes of maximization, ELBO as given in Eqn. (3) is thus equivalent to local entropy (upto an additive constant).

@@

In response to your previous comments, we have tested SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and can add the following results to the paper. After a hyper-parameter search, the best results we obtained yielded a test error of LeNet of 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs. Disregarding the slow convergence of SGLD, its generalization error is worse than the results in our paper, viz. 0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD). For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (and get 0.47% error with Santa). To our knowledge, there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.

[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).
-------------------------------------------------------------------------

You claim that "For modern deep networks with a few million weights, computing the Hessian-vector product as done in HS97 is prohibitive".  But in general the HVP can be computed for only twice the time cost as a gradient evaluation.  Can you elaborate on why this is expensive in your case?

The above comment is for HVP of the loss function of a general deep network, not local entropy. Indeed, HVPs can be computed using forward differencing with two back-props and Hessian-free methods do exploit this, although to a limited degree of success [M10]. Such an approximation is susceptible to numerical errors --- especially in high dimensions [P94, M10]. Moreover, one typically has to average the HVP over many samples in the dataset which is expensive (for instance, the authors in [LSP93] average over a few hundred samples, which roughly translates to 5-10x SGD _per_ HVP). More accurate algorithms like that of [P94] also require this averaging over samples.

[LSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS 1993.
[P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.
[M10] Martens, James. "Deep learning via Hessian-free optimization." ICML. 2010.

### 2017-01-02 doc/rebuttal.txt

Note that the above equation is about the entire Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn. 3. We can add this discussion to the paper.

[CFG14] Chen, Tianqi, Emily B. Fox, and Carlos Guestrin. "Stochastic Gradient Hamiltonian Monte Carlo." ICML. 2014.
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).

----------------------------------------------------------------------

Title: You cannot do SVI on the network weights unless you choose a “moving prior”, a feature of our scheme.

To address the objection 1) let \Xi denote the dataset, z  the weights and x the parameters of a variational distribution q_x(z). The ELBO can then be written as
(i) \log p(\Xi)
        \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z));
and maximized wrt x. The distribution p(z) is a fixed (parameter-free) prior, which one has to postulate.

On the other hand, Eqn. 4 in the paper can be used to write the log of local entropy, where f = -\logp(\Xi | z):
(ii) \log F(x,\gamma)
        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz.
We are unaware of any way to choose a prior p(z) and a variational family q_x(z) that that makes (i) resemble (ii), and therefore interpret our method as “integrating out […] the posterior over neural network parameters.” The only way we can do so is to pick a specific “prior" that depends on the parameter x (hence, not really a prior). For instance, one could choose a uniform variational family and a Gaussian “prior" with mean x to make (ii) resemble ELBO. In this case p(z) would not be fixed, but it would “move” along with the current iterate x. This "moving prior" is a crucial feature of our proposed algorithm.

Concerning the relation to SGLD, consider Belief Propagation (BP). Our proposed algorithm relates to the "focusing-Belief Propagation" variant (fBP), rather than the standard one [BBC16]. The difference between BP and fBP is analogous to that between SLGD and entropy-SGD: the latter operates on a transformation of the landscape of the former, exploiting local entropic effects. This difference is crucial and indeed related to the "moving prior" of the previous discussion; plain SGLD (or BP) can only trade accuracy for entropy via the temperature parameter, which does not allow to directly use any geometric information on the landscape and does not help with narrow minima.

In view of your comment 2), we also implemented SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and will add the following results to our paper: After a hyper-parameter search, the best rwe obtained were a test error of LeNet on 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs. Disregarding the slow convergence of SGLD, its generalization error is slightly worse than the results in our paper, viz. 0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD). For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (0.47% with Santa), there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.

[BBC16] Baldassi, Carlo, et al. "Unreasonable effectiveness of learning neural networks: From accessible states and robust ensembles to basic algorithmic schemes." PNAS (2016).
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).

-------------------------------------------------------------------------

Title: Re: Is the Hessian-vector product actually expensive?

Comment 3) above is for HVP of the loss function of a general deep network, not local entropy. Indeed, HVPs can be computed using forward differencing with two back-props and Hessian-free methods do exploit this, although to a limited degree of success [M10]. Such an approximation is susceptible to numerical errors --- especially in high dimensions [P94, M10]. One typically averages the HVP over many samples in the dataset which is expensive (for instance, the authors in [LSP93] average over a few hundred samples, which roughly translates to 5-10x one iteration of vanilla SGD). More accurate algorithms like that of [P94] also require this averaging over samples.

Practically speaking, one could argue that accuracy in the approximation of the HVP does not matter in SGD; what matters is only the local curvature at the scale of the perturbation, which needs to be chosen carefully.  In this case, indeed, approximate computation of HVP can be considered cheap.

[LSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS (1993).
[P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.
[M10] Martens, James. "Deep learning via Hessian-free optimization." ICML (2010).
